Highly Available Transactions: Virtues and Limitations

### 2013-03-31 writeup/intro.tex
more writing

Unfortunately, the relationship of the CAP Theorem and ACID semantics
has not been well explored. We have a strong understanding of weak
isolation in the single-server context from which it
originated~\cite{adya, ansicritique, gray-isolation} and many papers
offer techniques for providing distributed
serializability~\cite{bernstein-concurrency, spanner, granola,
  daudjee-session, krikellas-bargain, calvin} or snapshot
isolation~\cite{daudjee-snapshot,generalizedsnapshot, kemme-snapshot,
  walter}. The distributed computing, programming language, and
parallel hardware literature also contains many consistency models for
single operations on replicated objects~\cite{pnuts, herlihy-art,
  cops, eiger, cac, sessionguarantees}. However, the literature has
largely ignored semantic guarantees for multiple operations operating
on multiple data items in a highly available distributed environment.

>>>>

Unfortunately, the relationship of the CAP Theorem and ACID semantics
has not been well explored. We have a strong understanding of weak
isolation in the single-server context from which it
originated~\cite{adya, ansicritique, gray-isolation} and many papers
offer techniques for providing distributed
serializability~\cite{bernstein-concurrency, kemme-classification,
  spanner, granola, daudjee-session, krikellas-bargain, calvin} or
snapshot isolation~\cite{daudjee-snapshot,generalizedsnapshot,
  kemme-snapshot, walter}. The distributed computing, programming
language, and parallel hardware literature also contains many
consistency models for single operations on replicated
objects~\cite{pnuts, herlihy-art, cops, eiger, cac, sessionguarantees,
  ceri-mechanism, chen-mechanism}. However, the literature has largely
ignored semantic guarantees for multiple operations operating on
multiple data items in a highly available distributed environment.

### 2013-03-31 writeup/relatedwork.tex
more writing

Although we are not aware of any related work specifically addressing
what properties are available in a highly available system, there are
several systems and related research results upon which we draw in
this work.

Weak consistency and high availability have been well
studied. Serializability has long been known to be
unachievable~\cite{davidson-survey} and Brewer's CAP Theorem has
attracted considerable attention~\cite{gilbert-cap}. The research has
begun to offer hints regarding the connection between ``weak
consistency'' and low latency~\cite{abadi-pacelc}, while several
studies~\cite{adya, ansicritique} examine weak isolation guarantees,
grounded in a single-server context. The connection between weak
consistency models in a distributed setting~\cite{calm, crdt} and high
availability is relatively unexplored. With the exception of a
preliminary workshop paper on this work~\cite{hat-hotos}---upon which
this work builds, providing impossibility results, a full
taxonomization, experimental and application analysis (basically all
but Section 4.1)---we believe this paper is the first to explore this
area.

There has been a recent resurgence of interest in distributed
multi-object semantics, both in academia~\cite{kraska-s3, granola,
  gstore, mdcc, redblue, cops, eiger, walter,calvin, swift} and
industry~\cite{megastore, orleans, spanner}. As discussed in
Section~\ref{sec:motivation}, classic ACID databases provide strong
semantics but their lock-based and traditional multi-versioned
implementations are unavailable in the presence of
partitions~\cite{bernstein-concurrency, bernstein-book,
  gray-isolation}. Similar semantics are common in recent
work. Notably, Google's Spanner provides externally serializable
transactions. While Spanner fits Google's read-mostly workload well,
read/write transactions use two-phase locking (incurring at least one
round trip time over wide-area networks), are unavailable to clients
in a non-majority partition, and cannot read their
writes~\cite{spanner}. These latency and availability penalties are
fundamental to providing strong consistency. For users willing to
tolerate these costs, Spanner, or similar strongly consistent,
unavailable systems---including Calvin~\cite{calvin},
G-Store~\cite{gstore}, Gemini~\cite{redblue}, Granola~\cite{granola},
HBase~\cite{hbase}, MDCC~\cite{mdcc}, Megastore~\cite{megastore},
Orleans~\cite{orleans}, and Walter~\cite{walter}---are a reasonable
choice. With HATs, we seek an alternative set of semantics that are
still useful but do not violate requirements for high availability or
low latency. Recent systems proposals such as Swift~\cite{swift},
Eiger~\cite{eiger}, and Bolt-on Causal Consistency~\cite{bolton}
provide transactional causal consistency guarantees with varying
availability and represent a new class of (sticky) HAT systems.

Wisemann et al. propose a three parameter classification for one-copy
serializable database systems, performing a more extensive
classification than our overview in Setion XX, including eager
database replication~\cite{kemme-classification}. Additional, several
surveys discuss distributed mechanisms for achieving one-copy
serializability~\cite{wisemann-survey}, linearizability, consistent
snapshots, and recency bounds~\cite{ceri-mechanism, chen-mechanism}.

>>>>

Weak consistency and high availability have been well
studied. Serializability has long been known to be
unachievable~\cite{davidson-survey} and Brewer's CAP Theorem has
attracted considerable attention~\cite{gilbert-cap}. The research has
begun to offer hints regarding the connection between ``weak
consistency'' and low latency~\cite{abadi-pacelc}, while several
studies examine weak isolation guarantees~\cite{adya,
  ansicritique}. However, as we have discussed, the connection between
weak consistency models in a distributed setting~\cite{calm, crdt} and
high availability is relatively unexplored. With the exception of a
preliminary workshop paper introducing HATs~\cite{hat-hotos}--- which
this work expands, providing impossibility results, a full
taxonomization, and experimental and application analysis (basically
all but Section 4.1)---we believe this paper is the first to explore
this area. Nonetheless, several studies rigorously classify non-HAT
semantics: Wisemann et al. have proposed a three parameter
classification for one-copy serializable database systems, performing
a more extensive classification than our overview in
Section~\ref{sec:evaluation}, including eager database replication
techniques~\cite{kemme-classification}. Additionally several earlier
surveys discuss distributed mechanisms for achieving one-copy
serializability~\cite{wisemann-survey}, linearizability, consistent
snapshots, and recency bounds~\cite{ceri-mechanism, chen-mechanism}.

There has been a recent resurgence of interest in distributed
multi-object semantics, both in academia~\cite{kraska-s3, granola,
  gstore, mdcc, redblue, cops, eiger, walter,calvin, swift} and
industry~\cite{megastore, orleans, spanner}. As discussed in
Section~\ref{sec:modernacid}, classic ACID databases provide strong
semantics but their lock-based and traditional multi-versioned
implementations are unavailable in the presence of
partitions~\cite{bernstein-concurrency, bernstein-book,
  gray-isolation}. Notably, Google's Spanner provides strong one-copy
serializable transactions. While Spanner is highly specialized for
Google's read-heavy workload, it relies on two-phase commit and
two-phase locking for read/write transactions~\cite{spanner}. As we
have discussed, the latency penalties associated with this design
choice are fundamental to serializability. For users willing to
tolerate these costs, Spanner, or similar strongly consistent,
unavailable systems---including Calvin~\cite{calvin},
G-Store~\cite{gstore}, Gemini~\cite{redblue}, Generalized Snapshot
Isolation~\cite{generalizedsnapshot}, Granola~\cite{granola},
HBase~\cite{hbase}, MDCC~\cite{mdcc}, Megastore~\cite{megastore},
Orleans~\cite{orleans}, Postgres-R~\cite{kemme-thesis}, and
Walter~\cite{walter}---are a reasonable choice. We note that many
replication techniques target variants of Snapshot
Isolation~\cite{middleware-db}.

With HATs, we seek an alternative set of semantics that are
still useful but do not violate requirements for high availability or
low latency. Recent systems proposals such as Swift~\cite{swift},
Eiger~\cite{eiger}, and Bolt-on Causal Consistency~\cite{bolton}
provide transactional causal consistency guarantees with varying
availability and represent a new class of (sticky) HAT systems.

Middleware-based Database Replication: The Gaps Between Theory and Practice
Most uses Snapshot Isolation
